Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-07 ms.tex
conclusions 9

\item Including the effects of energy transport by electron conduction results in only moderate changes to the key properties of the flow (e.g. stagnation radius, mass accretion rate, and thermal stability) for causal values of the conduction saturation parameter $\phi < 0.1$ (eq.~[\ref{eq:conduction}]).

\item Radiative cooling has a more pronounced influence on the structure of the gas flow.  The approximate condition for cooling instabilities develop at all radii is that the radiative cooling rate exceed the external gas heating rate near the stagnation radius.  This condition is equivalent to the more commonly invoked $t_{\rm cool} \gg t_{\rm ff}$ criterion (e.g.~\citealt{McCourt+12}; \citealt{Li&Bryan14a}).

  \item The location of the stagnation radius, and hence the SMBH accretion rate, is a sensitive function of the gas heating rate $\propto v_w^{2}$.  Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the star formation history in the galactic nuclear region.  A recent star burst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for an exclusively old stellar population, mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation: energy input is usually dominated by young stars, while mass input is dominated by the old stellar population (Appendix \ref{app:windheat}).

Quantitatively, $\rs\propto v_w^{-2}$, implying that $\Mdot\propto  v_w^{-3.8 (-2.4)}$ for core (cusp) galaxies.  However, $\Mdot$ can increase even more rapidly with decreasing $v_w$ for two reasons: (1) when $v_w$ is comparable or less than the stellar velocity dispersion ($\zeta < \zeta_c$; eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar bulge and cannot produce a flow with a positive kinetic energy interior to the break radius of the stellar density profile (Appendix \ref{app:be}); (2) For low $v_w$, gas near the stagnation radius becomes thermally unstable, which based on previous numerical studies (e.g.,\citealt{Ciotti+10}) is found to result in time-dependent cycles of alternating high and low accretion states ($\S\ref{sec:cycle}$).

\item Galactic nuclei harboring low mass black holes $\Mbh \lesssim 10^8\Msun$ experiencing star formation histories to match their average derived from cosmological simulations receive stellar wind heating of $v_w\sim 800$ km s$^{-1}$ (Fig.~\ref{fig:vwSources}) sufficient to result in thermally-stable, steady-state accretion.  The resulting average Eddington-normalized accretion rate is predicted to increase weakly with $M_{\bullet}$ (Fig.~\ref{fig:miller}), in tension with the (also weak) downsizing trend of measured nuclear X-ray luminosities of early-type galaxies (e.g., \citealt{Miller+15}).  However, given the large uncertainties in translating predictions for the gas inflow rate on parsec scales to the disk X-ray luminosity, these differences may not be significant.

@@

\section*{Acknowledgments}
We acknowledge helpful conversations with Greg Bryan, Jules Halpern, Kathryn Johnston, Jerry Ostriker, and Eliot
Quataert.  This work relied on Columbia University's Yeti compute
cluster,and we acknowledge the Columbia HPC support staff for
assistance.  BDM gratefully acknowledges support from NASA {\it Fermi}
grant NNX14AQ68G, NSF grant AST-1410950, and the Alfred P. Sloan
Foundation.

>>>>

\item Including the effects of heat transport by electron conduction results in only moderate changes to the key properties of the flow (e.g. stagnation radius, mass accretion rate, and thermal stability) for causal values of the conduction saturation parameter $\phi < 0.1$ (eq.~[\ref{eq:conduction}]).

\item Radiation has a more pronounced influence on the structure of the gas flow.  Cooling instabilities are likely to develop when the radiative cooling rate exceed the rate of external heating near the stagnation radius.  This condition is equivalent to the more commonly invoked criterion $t_{\rm cool} \gg t_{\rm ff}$ criterion (e.g.~\citealt{McCourt+12}; \citealt{Li&Bryan14a}).

  \item The location of the stagnation radius, and hence the SMBH accretion rate, is a sensitive function of the gas heating rate $\propto v_w^{2}$.  Quantitatively, $\rs\propto v_w^{-2}$, implying that $\Mdot\propto  v_w^{-3.8 (-2.4)}$ for core (cusp) galaxies.  However, $\Mdot$ can increase even more rapidly with decreasing $v_w$ for two reasons: (1) when $v_w$ is comparable or less than the stellar velocity dispersion ($\zeta < \zeta_c$; eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar bulge and cannot produce a flow with a positive kinetic energy interior to the break radius of the stellar density profile (Appendix \ref{app:be}); (2) For low $v_w$, gas near the stagnation radius becomes thermally unstable, which based on previous numerical studies (e.g.,\citealt{Ciotti+10}) is found to result in time-dependent cycles of alternating high and low accretion states ($\S\ref{sec:cycle}$).

\item Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the star formation history in the galactic nuclear region.  A recent star burst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for an exclusively old stellar population, mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation: energy input is usually dominated by young stars, while mass input is dominated by the old stellar population (Appendix \ref{app:windheat}).  Galactic nuclei harboring low mass black holes $\Mbh \lesssim 10^8\Msun$ experiencing star formation histories to match their average derived from cosmological simulations receive stellar wind heating of $v_w\sim 800$ km s$^{-1}$.

\item Type Ia SNe only provide a continuous heating source exterior to the radius $\rIa$ (eq.~[\ref{eq:RIa}]) where the time interval between subsequent SNe exceeds the local dynamical timescale.  The heating provided by the average star formation history results in $\rIa < \rs$, except for the most massive galaxies.  Nevertheless, even if other sources of heating are low enough that $\rIa > \rs$, the stagnation radius is unlikely to exceed $\rIa$ due to Ia heating.  SN Ia thus cap the SMBH accretion rate at the value $\dot{M}_{Ia}$ (eq.~[\ref{eq:MdotIa}]), except possibly in high mass elliptical galaxies with large break radii, for which $\zeta(v_w^{\rm Ia}) \gtrsim \zeta_c$.

\item Stellar wind heating due to the average star formation history is sufficient to result in thermally-stable, steady-state accretion.  The resulting average Eddington-normalized accretion rate is predicted to increase weakly with $M_{\bullet}$ (Fig.~\ref{fig:miller}), in tension with the (also weak) downsizing trend of measured nuclear X-ray luminosities of early-type galaxies (e.g., \citealt{Miller+15}).  However, given the large uncertainties in translating predictions for the gas inflow rate on parsec scales to the disk X-ray luminosity, these differences may not be significant.

@@

\section*{Acknowledgments}

We acknowledge helpful conversations with Greg Bryan, Jules Halpern,
Kathryn Johnston, Jerry Ostriker, and Eliot Quataert.  This work
relied on Columbia University's Yeti compute cluster,and we
acknowledge the Columbia HPC support staff for assistance.  BDM
gratefully acknowledges support from NASA {\it Fermi} grant
NNX14AQ68G, NSF grant AST-1410950, and the Alfred P. Sloan Foundation.
